---
title: "HW1 Part 2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 5
### a) Function to simulate y values from linear model

```{r get_y}
gety <- function(x) {
  y <- 2*x^2 + rnorm(length(x),0,2)
  return(y)
}
```

### b) Simulation

```{r sim}
##run simulation 100000 times
reps <- 100000

##create an array to store the estimated slope from each rep
store.y0 <- array(0, reps)
store.f_hat0 <- array(0, reps)

##time loop
start_time <- Sys.time() ##start time of loop

set.seed(4630)

for (i in 1:reps)

{
  ##generate the values of x
  x <- rep(seq(1,10,1),20)
  
  ##simulate values of y
  y <- gety(x) 
  
  ##use least squares to obtain regression equation on simulated data
  result <- lm(y~x)

  ##store the important values from this rep
  store.y0[i] <- gety(c(7)) #y0
  store.f_hat0[i] <- predict(result, data.frame(x=c(7))) #y_hat0
  
  #track progress
  if(i %% 10000 == 0) print(paste("Iteration", i))
}

end_time <- Sys.time() ## end time of loop
end_time - start_time ##time taken by loop
```

### c) Calculate the expected test MSE at x0 = 7

```{r testMSE}
exp_test_MSE <- (1/reps)*sum((store.y0-store.f_hat0)^2)
exp_test_MSE
```

### d) Calculate f_bar(x0) at x0 = 7

```{r}
f_bar_x0 <- mean(store.f_hat0)
f_bar_x0
```

### e) Using your 100,000 values of y0 and f_hat(x0), calculate each of the three sources of error, report each of their values, and then add them up (and report the value when added up).

```{r}
var_f_hat <- mean((store.f_hat0-f_bar_x0)^2)
var_f_hat

bias_squared <- (98-f_bar_x0)^2
bias_squared

var_e <- mean((store.y0-98)^2)
var_e

var_f_hat + bias_squared + var_e
```

### f) The third source of error, E[(y0 - f(x0))^2], should be close to 4 (it should be 4, theoretically). In one sentence, briefly explain why.

It should be 4 because that is the variance of the standard normal distribution from which we generated the error terms in the line: rnorm(length(x),0,2).

### g) Based on your values from 5c and 5e, what is the difference between the LHS and RHS of (1)?

```{r}
exp_test_MSE - (var_f_hat + bias_squared + var_e)
```

### h) Be sure to include the names of classmates you worked with on this question.

Grady Bartro
