---
title: "Lab 4c Quiz"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(199)
library(faraway)
library(MASS)
library(klaR)
library(ICS)
```

## 0

```{r}
wcgs_interest <- wcgs[,c("age", "sdp", "dbp", "cigs", "chd")]

##create training and test data
sample.data<-sample.int(nrow(wcgs_interest), floor(.50*nrow(wcgs_interest)), replace = F)
train<-wcgs_interest[sample.data, ]
test<-wcgs_interest[-sample.data, ]
```

## 1

Because it is qualitative

## 2

```{r}
##subset dataframe by gender
chd_yes<-train[which(train$chd=="yes"),]
chd_no<-train[which(train$chd=="no"),]

##MVN tests for females
ICS::mvnorm.kur.test(chd_yes[,1:4])
ICS::mvnorm.skew.test(chd_yes[,1:4])

##MVN tests for males
ICS::mvnorm.kur.test(chd_no[,1:4])
ICS::mvnorm.skew.test(chd_no[,1:4])
```

Probably don't mean much if our goal is classification, not esimating posterior probabilities.

## 3

```{r}
##check dummy coding for gender
contrasts(train$chd)

##Carry out LDA on training data
lda.chd <- MASS::lda(chd ~ ., data=train)
##obtain ouput from LDA
lda.chd
```

All predictors are positive, so older people, people with higher blood pressure, and people who smoke more will be classified as having chd. The best predictor is age.

## 4

```{r}
##boundaries based on 2 predictors
klaR::partimat(chd ~ ., nplots.vert=2, nplots.hor=3, data=train, method="lda")
```

It takes a very specific and unlikely combination of traits to be classified as having chd.

## 5

```{r}
##predictions on training data. 
lda.train <- predict(lda.chd)
##Confusion matrix on training data. Rows represent actual value, cols represent pred value
table(train$chd, lda.train$class)

##accuracy on training data
mean(train$chd == lda.train$class)

##predictions on test data. 
lda.test <- predict(lda.chd,test)
##confusion matrix. By default, threshold is 0.5
table(test$chd,lda.test$class)

##accuracy on test data
mean(test$chd == lda.test$class)

##posterior probabilities for first 6 observations of test data
head(lda.test$posterior)

##threshold of 0.2
table(test$chd, lda.test$posterior[,2]>0.2)

##ROC and AUC
library(ROCR)
preds<-lda.test$posterior[,2]
rates<-ROCR::prediction(preds, test$chd)
roc_result<-ROCR::performance(rates,measure="tpr", x.measure="fpr")
plot(roc_result, main="ROC Curve for Whether a Person Has CHD")
lines(x = c(0,1), y = c(0,1), col="red")
```

## 6

```{r}
auc<-ROCR::performance(rates, measure = "auc")
auc@y.values
```

## 7

```{r}
confusion.mat<-table(test$chd,preds > 0.07)
confusion.mat

accuracy <- (891+89)/(891+89+549+48)
accuracy

fpr <- 549/(549+891)
fpr

fnr <- 48/(48+89)
fnr
```

## 8

```{r}
##Use QDA on training data
qda.chd <- MASS::qda(chd ~ ., train)
qda.chd

##Quadratic boundaries based on 2 predictors
klaR::partimat(chd ~ ., data=train, method="qda")
```

## 9

```{r}
##predictions on test data
qda.test <- predict(qda.chd,test)
table(test$chd,qda.test$class)

##ROC and AUC
preds.qda<-qda.test$posterior[,2]
rates.qda<-ROCR::prediction(preds.qda, test$chd)
roc_result.qda<-ROCR::performance(rates.qda,measure="tpr", x.measure="fpr")
plot(roc_result.qda, main="ROC Curve for Whether a Person Has CHD")
lines(x = c(0,1), y = c(0,1), col="red")
auc<-ROCR::performance(rates.qda, measure = "auc")
auc@y.values
```

## 10

```{r}
confusion.mat<-table(test$chd,preds.qda > 0.07)
confusion.mat

accuracy <- (932+83)/(932+83+508+54)
accuracy

fpr <- 508/(508+932)
fpr

fnr <- 54/(54+83)
fnr
```

## 11

LDA (lower accuracy but also lower false negative rate)

## 12

LDA (higher accuracy than logistic while maintaining false negative rate)
